---
title: "Group13_DataStory"
author: "Marti Garriga Marsans, Yihua Chan, Saumya Srivastava, Thomas Tyteca, Yanjun He, Zongqiang Lian"
date: "2023-11-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# 1. Importing libraries
```{r}
library(gapminder)
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(lubridate)
library(showtext)
library(janitor)
library(vroom)
library(ggtext)

font_add_google("Montserrat", "Montserrat")
font_add_google("Ubuntu", "Ubuntu")
font_add_google("Oswald", "Oswald")
font_add_google("Lato", "Lato")
```


# 2. Importing datasets
These datasets have been cleaned beforehand.
```{r, warning = FALSE}
unzip("df_dom.csv.zip")
club_results <- read_csv("df_dom.csv") # Domestic (club) matches around the world

english_club_loss <- read_csv("eng_dom_loss.csv") # Domestic matches where English club loses

english_club_win <- read_csv("eng_dom_win.csv") # Domestic matches where English club wins

fin_UCL <- read_csv("fin_UCL.csv") # Champions League Matches

finance_final <- read_csv("finance_final.csv") %>% clean_names() # FTSE100 index 2000 - 2023

final_overall <- read_csv("final_overall.csv") %>% 
  distinct(date, home, away, home_team, away_team, .keep_all = TRUE) # International (national team) matches joined with FTSE100 financial information

int_games <- read_csv("int_games.csv") # International (national team) matches
```

# 3. Plot on the evolution of FTSE100 Index
```{r}
ftse100_evolution_plot <- finance_final %>% 
  ggplot() +
  geom_line(aes(x = date, y = open), colour = "darkblue", size = 0.75) +
  theme_minimal() +
  labs(title = 
      "<b>Evolution of the <span style='color:darkblue'>FTSE100 index</span> since 2001</b>",
      x = NULL,
      y = NULL) +
  theme(plot.title = element_textbox_simple(size = 14, colour = "grey15", margin = margin(t = 10, b = 10)),
        axis.text.x = element_text(hjust = 0.5),
        axis.ticks.x = element_line(color = "grey15"),
        text = element_text(size = 12, family = "Montserrat")) +
  theme(plot.margin = margin(r = 10, l = 10))

ftse100_evolution_plot
```
  
# 4. Key dates in English national team history
And how the stock price reacted to the outcomes the next day.
```{r}
date_int <- c('2018-07-09', '2021-07-12', '2018-07-11', '2016-06-27', '2012-06-26', '2010-06-29')
game_int <- c("Advances to World Cup semi-finals", "Loses EuroCup final (hosted in Wembley)", "Eliminated from 2018 WC", "Eliminated from 2016 EuroCup", "Eliminated from 2012 EuroCup", "Frank Lampard's Ghost Goal that led to 2010 WC elimination")
date_int <- as.Date(date_int)

df_int_key <- data.frame(date = date_int, game = game_int)

df_int_key_finance <- df_int_key %>% 
  left_join(select(finance_final, date, change_percent), by = "date")

avg_daily_change <- data.frame(date = '2023-12-31', game = "Average daily change", change_percent = mean(finance_final$change_percent))

df_int_key_finance <- rbind(df_int_key_finance, avg_daily_change)

int_key_finance_plot <- ggplot(df_int_key_finance, aes(x = reorder(str_wrap(game, width = 10), -change_percent), y = change_percent)) +
  geom_bar(stat = "identity",
           fill = ifelse(df_int_key_finance$change_percent > 0, "darkgreen", "tomato")) +
  geom_text(aes(label = paste0(round(change_percent, 2), "%")),
            vjust = ifelse(df_int_key_finance$change_percent > 0.5, 1.25, -0.5),
            colour = ifelse(abs(df_int_key_finance$change_percent) > 0.5, "white", "grey15"),
            fontface = "bold", size = 5) +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        axis.text.y = element_blank(),
        plot.title = element_text(face = "bold"),
        text = element_text(family = "Montserrat"),
        axis.text.x = element_text(face = ifelse(df_int_key_finance$game == "Eliminated from 2018 WC", "bold", "plain"))) +
  labs(title = "How did the stock market react to these historical matches?",
       subtitle = "% change in FTSE100 price on the following day",
       x = NULL,
       y = NULL)

int_key_finance_plot
```

# 5. Elimination from 2010 WC - Controversial unallowed goal
```{r}
finance_2010_elimination <- finance_final %>% 
  filter(date > '2010-06-10', date < '2010-07-10')

finance_2010_elimination_month <- ggplot(finance_2010_elimination, aes(x = date)) +
  geom_line(aes(y = change_percent), colour = "darkblue", size = 1.5) +
  labs(title = "So, what happened after the game?",
       subtitle = "Zooming out on Summer 2010...",
       x = NULL,
       y = "% Change in FTSE100") +
  geom_vline(xintercept = as.Date('2010-06-28'), colour = "tomato", linetype = "dashed") +
  annotate("text", x = as.Date('2010-06-28'), y = 2.5, label = "England kicked out of the World Cup", vjust = -0.5, hjust = 1.05, colour = "tomato", family = "Montserrat") +
  scale_y_continuous(labels = function(y) paste0(y, "%")) +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        text = element_text(family = "Montserrat"),
        plot.title = element_text(face = "bold"))

finance_2010_elimination_month
```

# 6. Linear regression model to predict stock movements
```{r}
england_matches <- int_games %>%
  filter(home_team == 'England' | away_team == 'England')

# Determine the winner of each match
england_matches <- england_matches %>%
  mutate(winner = case_when(
    home_score > away_score & home_team == 'England' | away_score > home_score & away_team == 'England' ~ 'England',
    home_score == away_score ~ 'Draw',
    TRUE ~ 'Opponent'
  ))

# Convert date to Year and filter for 2001 to 2019
england_matches <- england_matches %>%
  mutate(year = as.numeric(format(as.Date(date, "%Y-%m-%d"), "%Y"))) %>%
  filter(year >= 2001, year <= 2019)

finance_final$date <- as.Date(finance_final$date)

# Join the datasets on the date column
combined_data_england <- left_join(england_matches, finance_final, by = "date") 
# View the first few rows of the combined dataset
length(combined_data_england)

combined_data_england$date <- ymd(combined_data_england$date) # ymd stands for "year-month-day"
# Extracting the day
combined_data_england$day <- day(combined_data_england$date)

# Extracting the month
combined_data_england$month <- month(combined_data_england$date)

# Fitting the model
model <- lm(price~  winner + tournament + month + home_team + away_team, data = combined_data_england)

# Summary of the model
summary(model)
```

# 7. Relationship between wins in a period and the average stock price within it
```{r, warning=FALSE}
combined_data_england_2 <- combined_data_england %>%
  mutate(eng_win = ifelse(winner == "England", 1, 0))

temp_df_eng <- combined_data_england_2 %>% 
  group_by(year,month) %>%
  filter(eng_win==1, year>2010) %>% 
  summarise(
    wins = n(),
    average_price = mean(price, na.rm = TRUE))

temp_df_eng %>% 
  filter(wins >= 2.1) %>% 
  ggplot(aes(x = wins, y = average_price)) +
  geom_point(color = "dark blue", size = 3) +
  geom_smooth(method = 'lm', se = FALSE, color = "tomato", linetype = "dashed") +
  theme_minimal() +
  labs(
    title = "English national team wins vs. Average price of FTSE100",
    subtitle = "An outstanding season from the Three Lions can mean excellent returns for FTSE100",
    x = "Number of Wins",
    y = "Average Price"
  ) +
  scale_x_continuous(breaks = seq(min(temp_df_eng$wins), max(temp_df_eng$wins), by = 2)) +
  scale_y_continuous(breaks = seq(5000, 8000, by = 500)) +
  theme(
    text = element_text(family = "Montserrat"),  # Set the global font family
    plot.title = element_text(face = "bold", margin = margin(b = -1, unit = "pt")),
    plot.subtitle = element_text(hjust = 0, margin = margin(t = 1, unit = "pt")),
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold"),
    axis.text = element_text(color = "grey15"),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(color = "gray", linetype = "dotted")
  )
```

# 8. Win percentage of the English national team over the years
```{r}
winning_teams <- int_games %>% 
  distinct(date, home_team, away_team, .keep_all = TRUE) %>%
  group_by(date, home_team, away_team) %>% 
  summarise(home_score, away_score) %>% 
  mutate(winning_team = if_else(home_score > away_score, home_team, if_else(home_score < away_score, away_team, "Draw")))

win_percentage <- winning_teams %>% 
  group_by(year = year(date)) %>% 
  filter(year > 2000) %>% 
  summarise(total_games = n(), 
            total_wins = sum(winning_team == "England"), 
            win_percentage = (total_wins/total_games)*100) %>% 
  mutate(average = mean(win_percentage))

win_perc_plot <- win_percentage %>% 
  ggplot() +
  geom_line(aes(x = year, y = win_percentage), colour = "darkblue", size = 1) +
  theme_minimal() +
  labs(title = 
      "<b>The Three Lions have <span style='color:tomato'>won over 60% of their games</span> since 2001</b><br>
      <span style='font-size:12pt;color:darkblue'>Yearly win percentage of the English national team between 2001 and 2023</span>",
      x = NULL,
      y = NULL) +
  theme(plot.title = element_textbox_simple(size = 16, colour = "grey15", margin = margin(t = 10, b = 10)),
        axis.text.x = element_text(hjust = 0.5),
        axis.ticks.x = element_line(color = "grey15"),
        text = element_text(size = 12, family = "Montserrat")) +
  geom_hline(aes(yintercept = average), linetype = "dotted", color = "tomato", size = 1) +
  geom_text(data = data.frame(x = 2023, y = 63.5, label = "61.27%"),
            aes(x = x, y = y, label = label),
            family="Montserrat",
            colour = "tomato",
            size = 4,
            hjust = 0,
            lineheight = .8,
            inherit.aes = FALSE,
            fontface = "bold") +
  lims(x = c(2001, 2024)) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme(plot.margin = margin(r = 10, l = 10))

win_perc_plot
```


